Cover Times of Random Walks

### 2019-04-28 paper.tex
yay empirical!

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{varioref}
\usepackage{verbatim} 
\usepackage{multicol}
\usepackage{enumerate}
\usepackage[normalem]{ulem}
%\usepackage[margin=1in]{geometry}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{thm-restate}
\usepackage{enumitem}
\usepackage{upgreek}
%\usepackage{cite}
\usepackage{cite}

@@

There are many connections between electrical networks
and random walks \cite{DS84}.
Physical properties like voltage, current, and resistance
all have probabilistic interpretations.
In this section, we explore electrical networks to build intuition about 
random walks.

@@

The current between two vertices is the product of the
difference in their voltages and the conductance of their edge.
We may think of current as the net number of movements through an edge.
Kirchoff's Current Law says that the sum of current on any vertex is 0.
Probabilistically, of course every time we enter a vertex we must also leave it.
Using Kirchoff's Current Law and the definition of current,
we justify the probabilistic interpretation of voltage
\begin{align}
\sum_y i_{x,y} &= 0 = \sum_y (v_x - v_y)C_{x,y} \nonumber \\
v_x \sum_y C_{x,y} &= \sum_y C_{x,y} v_y \nonumber \\
v_x &= \sum_y \frac{C_{x,y}}{C_x} v_y = \sum_y \pr_{x,y} v_y. \nonumber
\end{align}
We can probabilistically verify that $v_x = \sum_y \pr_{x,y} v_y$
using first step analysis on the probability of reaching $a$ before $b$.

@@

@@

If we let the label of each graph in \cref{fig:example}
denote the cover time from that case, we have the following system
of equations:
\begin{align}
\tcov &= 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{b} \nonumber \\
\textrm{a} &= 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{c} \nonumber \\
\textrm{b} &= 1 + \frac{1}{2}\textrm{b} + \frac{1}{2}\textrm{e} \nonumber \\
\textrm{c} &= 1 + 1\textrm{e} \nonumber\\
\textrm{d} &= 1 + \frac{1}{3}\textrm{a} + \frac{1}{3}\textrm{e} + \frac{1}{3}\textrm{f} \nonumber \\
\textrm{e} &= 1 + \frac{1}{3}\textrm{c} + \frac{1}{3}\textrm{e} + \frac{1}{3}\dot 0\nonumber \\
\textrm{f} &= 1 + \frac{1}{2}\textrm{g} + \frac{1}{2}\dot 0\nonumber \\
\textrm{g} &= 1 + \frac{2}{3}\textrm{f} + \frac{1}{3}\dot 0\nonumber
\end{align}

@@

\subsection{Commute Time Identity} \label{sec:commute}
We use the electrical network analysis in \cref{sec:electric} to 
prove the commute time identity so we may use it to help us bound
cover times.

@@

\section{Cover Time Distribution}
\subsection{Complete Graph}
\cite{Du11}
We consider the distribution of the cover time of a random walk
on the complete graph with self edges.
Call $r$ the number of steps since we began the random walk.
So that the notation works out more easily,
set $r=1$ when we begin our walk at $X_0$.
Let $A_i$ be the event that vertex $i$ is not visited and
$N_n$ the number of unvisited vertices.
Then
\begin{align}
P(A_i) = (1-1/n)^r \nonumber
\end{align}
and
\begin{align}
\E[N_n] = n(1-1/n)^r. \nonumber
\end{align}

@@

Using the fact that $\cov \leq m$ if and only if the first $m$ steps
of the random walk visit all $n$ vertices
and $ne^{-r/n} \rightarrow e^{-x}$ as the assumption of \cref{thm:pois},
\begin{align}
\pr(\cov \leq r) &\rightarrow e^{-e^{-x}} \nonumber \\
\pr(\cov - n \log n\leq nx) &\rightarrow e^{-e^{-x}}. \nonumber
\end{align}

\subsection{Computation}
\todo{example of probability of birthday cover time in x=2190 vs x=1825 steps}

\todo{simulate cover time a thousand times on complete graph with n = 365}

\cite{Bo98}
\cite{Gr10}

@@

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{varioref}
\usepackage{verbatim} 
\usepackage{multicol}
\usepackage{enumerate}
\usepackage[normalem]{ulem}
%\usepackage[margin=1in]{geometry}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{thm-restate}
\usepackage{enumitem}
\usepackage{upgreek}
%\usepackage{cite}
\usepackage{cite}
\usepackage{listings}
\usepackage[toc, page]{appendix}

@@

There are many connections between electrical networks
and random walks \cite{Bo98}.
Physical properties like voltage, current, and resistance
all have probabilistic interpretations.
In this section, we explore electrical networks to build intuition about 
random walks.

@@

The current between two vertices is the product of the
difference in their voltages and the conductance of their edge.
We may think of current as the net number of movements through an edge.
Kirchoff's Current Law says that the sum of current on any vertex is 0.
Probabilistically, of course every time we enter a vertex we must also leave it.
Using Kirchoff's Current Law and the definition of current,
we justify the probabilistic interpretation of voltage
\begin{align}
\sum_y i_{x,y} &= 0 = \sum_y (v_x - v_y)C_{x,y} \nonumber \\
v_x \sum_y C_{x,y} &= \sum_y C_{x,y} v_y \nonumber \\
v_x &= \sum_y \frac{C_{x,y}}{C_x} v_y = \sum_y \pr_{x,y} v_y. \nonumber
\end{align}
We can verify that $v_x = \sum_y \pr_{x,y} v_y$
using first step analysis on the probability of reaching $a$ before $b$.

@@

Rely on \cite{BH94}
Always possible to find cover time but it gets hard as the graph gets bigger
Structure helps:
Complete graph,
Line graph,
Simple cycle (extended line graph)

In lieu of structure, we want to work with arbitrary graphs
Rely on \cite{DS84} for commute time relationship
Rely on \cite{LP17}
Matthews method is intuitive approach that uses same ideas as complete graph
Matthews method applied to binary tree with commute time relationship
Use \cite{AF14}
Explore the idea of a spanning tree and use it to bound time with commute identity

So far we've only worked with expected time
\cite{Du11}
Explore what happens for the variable and it's distribution

@@

If we let the label of each graph in \cref{fig:example}
denote the cover time from that case, we have the following system
of equations:
\begin{align}
&\tcov = 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{b}
&\textrm{a} = 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{c} \nonumber \\
&\textrm{b} = 1 + \frac{1}{2}\textrm{b} + \frac{1}{2}\textrm{e}
&\textrm{c} = 1 + 1\textrm{e} \nonumber \\
&\textrm{d} = 1 + \frac{1}{3}\textrm{a} + \frac{1}{3}\textrm{e} + \frac{1}{3}\textrm{f}
&\textrm{e} = 1 + \frac{1}{3}\textrm{c} + \frac{1}{3}\textrm{e} + \frac{1}{3}\dot 0\nonumber \\
&\textrm{f} = 1 + \frac{1}{2}\textrm{g} + \frac{1}{2}\dot 0
&\textrm{g} = 1 + \frac{2}{3}\textrm{f} + \frac{1}{3}\dot 0\nonumber \\
\end{align}

@@

\subsection{Commute Time Identity} \label{sec:commute}
\cite{DS84} We use the electrical network analysis in \cref{sec:electric} to 
prove the commute time identity so we may use it to help us bound
cover times.

@@

\section{Cover Time Distributions}
\subsection{Complete Graph}
\cite{Du11}
We consider the distribution of the cover time of a random walk
on the complete graph with self edges.
Call $r$ the number of steps since we began the random walk.
So that the notation works out more easily,
set $r=1$ when we begin our walk at $X_0$.
Let $A_i$ be the event that vertex $i$ is not visited and
$N_n$ the number of unvisited vertices.
Then
\begin{align}
P(A_i) = (1-1/n)^r \nonumber
\end{align}
and
\begin{align}
\E[N_n] = n(1-1/n)^r. \nonumber
\end{align}

@@

Using the fact that $\cov \leq m$ if and only if the first $m$ steps
of the random walk visit all $n$ vertices
and $ne^{-r/n} \rightarrow e^{-x}$ as the assumption of \cref{thm:pois},
\begin{align}
\pr(\cov \leq r) &\rightarrow e^{-e^{-x}} \nonumber \\
\pr(\cov - n \log n\leq nx) &\rightarrow e^{-e^{-x}}. \label{eqn:dist}
\end{align}

\subsection{The Birthday Problem}
Consider a complete graph (with self-edges) and 365 vertices.
We may think of the cover time variable of this graph as the number of people
until every birthday is represented.
We call this the birthday problem.

We use \cref{eqn:dist} to calculate the probability
that all birthdays are represented in populations of 
size $r=2190$ and $r=1825$, respectively.
The number of vertices $n$ is 365 so $n\log n = 2153$.
It follows that
\begin{align}
\pr(\cov \leq 2190) &= \pr \left(\frac{\cov - 2153}{365} \leq
\frac{37}{365}\right) \nonumber \\
&\approx e^{-e^{-.10137}} \approx 0.40511 \nonumber
\end{align}
but 
\begin{align}
\pr(\cov \leq 1825) &= \pr \left(\frac{\cov - 2153}{365} \leq
\frac{328}{365}\right) \nonumber \\
&\approx e^{-e^{-.8.9863}} \approx 0.08576. \nonumber
\end{align}

There appears to be a critical stage where it quickly becomes very likely
that all birthdays have been represented.
\cref{fig:birthday} shows the cumulative distribution of the birthday problem.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[scale=.5]
\node (graph) at (0,0) {\includegraphics[scale=.25, height=10cm]{birthdays}};
\node (title) at (1, 10) {Cover Times of the Birthday Problem (1000 trials)};
\node[rotate=90] (ylabel) at (-13,2) {$\pr(\cov \leq r)$};
\node (xlabel) at (0,-9) {$r$};
\end{tikzpicture}
\caption{The cumulative distribution of the birthday problem. The red line denotes the empirical distribution from 1000 trials. The blue line denotes the distribution from \cref{eqn:dist}.}\label{fig:birthday}
\end{figure}

%\cite{Gr10}

@@

\newpage
\begin{appendices}
\subsection{Birthday Problem}
\lstinputlisting{computation.R}
\end{appendices}
